


UniZ is a university administration platform built specifically for RGUKT students, faculty, and staff. Whether you need to check your grades, request an outpass, approve a student's leave request, or publish a campus announcement — UniZ brings everything into one place, accessible from any device.

The platform is organized around the roles that make up campus life. Students get a personal dashboard for academics and mobility requests. Faculty can upload grades and attendance. Wardens, caretakers, security staff, and administrators each have dedicated workflows that keep campus operations running smoothly. Every action flows through a structured, role-based system so the right people see the right information.

UniZ is a Progressive Web App (PWA), which means you can install it directly on your phone, tablet, or desktop from your browser — no app store required. Once installed, it works like a native app with a fast, focused experience tailored to your role.


\begin{tcolorbox}[colback=gray!5,colframe=primary,title=Documentation Note]

  
\begin{tcolorbox}[colback=gray!5,colframe=primary,title=Documentation Note]

    Get the app installed and log in to your portal in a few minutes.
  
\end{tcolorbox}

  
\begin{tcolorbox}[colback=gray!5,colframe=primary,title=Documentation Note]

    See what each role can do and find the features built for you.
  
\end{tcolorbox}

  
\begin{tcolorbox}[colback=gray!5,colframe=primary,title=Documentation Note]

    Sign in as a student and access your academic dashboard.
  
\end{tcolorbox}

  
\begin{tcolorbox}[colback=gray!5,colframe=primary,title=Documentation Note]

    Explore the authentication API for the UniZ platform.
  
\end{tcolorbox}

</CardGroup>


\section{Deep Technical Analysis}
The underlying system architecture for this module is built upon the \textbf{UniZ Microservices Framework}. 
This design ensures that each functional unit (Auth, Profile, Academics) remains isolated, preventing cascading failures across the university ecosystem.

\subsection{Operational Hardening}
Deployments are managed via K3s (Kubernetes) and containerized using high-performance Alpine Linux Docker images. 
Resource management is critical; for instance, the documentation service itself is provisioned with 2GB of RAM to ensure the responsive rendering of complex documentation graphs and state-management components.

\subsection{Enterprise Security Topology}
Every request entering the UniZ gateway is subjected to an aggressive JWT validation layer. 
Permissions are granular, mapping directly onto the RBAC (Role-Based Access Control) matrix defined in the core system specifications.

\subsection{Data Governance}
Prisma-driven data access ensures type safety and predictable query performance. 
We utilize PostgreSQL connection pooling to handle concurrent requests from the student portal and administrative dashboards simultaneously.
